BareTag Tool-Tracker\\ {\footnotesize {Team 3 - University of Massachusetts Amherst Senior Design Project 2025 }}

### 2024-12-19 PDR/PDR_draft_of_SDP_report/draft.tex
fic

\begin{abstract}
    Over the past 10 years, the majority of tools on a construction 
    site have converted from wired to battery-powered. While this 
    makes tools easy to move around, it also makes them easy to misplace
    and an easier target for theives. Data indicates that the construction
    idustry suffered nearly \$1,000,000,000 in loses due to tool theft in
    2016 alone [1], strongly indicating the need for a robust and effective 
    theft mitigation system. We propose the BareTag Tool Tracker, a novel
    approach to tool tracking that utilizes Ultra-Wideband(UWB) and Bluetooth 
    Low-Energy(BLE) radio in order to real-time track tools, materials, or other 
    valuable items on a construction site. The system utilizes a series
    of pre-placed Anchor posts, that send UWB pings to a Tag that is connected
    to a tool. Each Anchor can then calculate its distance to a Tag, relaying
    that information to a Base station over Long Range(LoRa) radio. The 
    base station runs the aggregated distance data through a multilateration
    algorithm that can calculate the Tag's location with $\pm$ 10 cm accuracy.
    The calculated location is then output to a local terminal, as well as 
    uploaded to a cloud database for future reference. Altogether, the BareTag Tool-Tracker
    is highly accurate ($\pm$ 10 cm), low-power (1 year of battery life), and
    scalable (increase range by adding additional Anchors).
\end{abstract}

@@

The specifications listed in Table 1 demonstrate the goals that our project 
aims to achieve by the time that we have a finalized product.
\begin{center}
\begin{table}
\caption{Design Goals and Testing Plan}
\begin{tabular}{|p{4cm}|p{4cm}|}
    \hline
    \textbf{Specification}  & \textbf{Testing Plan}\\
    \hline
    The system will track an item’s geo-coordinates with a reliability of 
    95\% within a specified 100 m x 100 m zone. 
    &
    Ping the Tag at randomized coordinates within the Sustainable 
    Engineering construction site.\\
    \hline
    The system will have less than 50 cm precision for tracking within 
    the specified zone.
    &
    Compare Tag’s measured location to its actual location 
    within a predefined grid.\\
    \hline
    The system will operate 24/7 for ~1 year without 
    replacing the batteries.
    &
    Measure the power consumption of the system in multiple modes.\\
    \hline
    Ensure that the Tag can be tracked via the Find My network outside 
    of the UWB zone.
    &
    Confirm tracking accuracy in scenarios with different saturations 
    of Find My devices by comparing with AirTags in the same location.\\
    \hline
    Should easily be scalable to increase the range of the UWB perimeter 
    by adding more Anchors at desired locations.
    &
    Demonstrate that a Tag’s trackable range in the UWB zone increases 
    when adding a new Anchor.\\
    \hline
\end{tabular}
\end{table}
\end{center}

@@

This device can be whatever system a supervisor desires to use as long as 
it can support a USB/Serial connection, or do so using adapters. The Host Device
performs the multilateration algorithms based on the data it receives from the LoRa Dongle.
The calculations get backhauled to a cloud database 
where the history of Tag locations can be mapped out to determine 
daily paths. This information is also displayed locally on a map of 
the desired area. The Host Device is also running the macless-haystack 
server, which acts as a means for tracking the Tags that have 
left the UWB area and are using BLE to integrate with the Apple Find 
My network. The Tags all have unique IDs that can be linked to tools or 
equipment, making it easy to parse the data and determine what items have 
been in what location.

>>>>

\begin{abstract}
    Over the past 10 years, the majority of tools on a construction 
    site have converted from wired to battery-powered. While this 
    makes tools easy to move around, it also makes them easy to misplace
    and an easier target for theives. Data indicates that the construction
    idustry suffered nearly \$1,000,000,000 in loses due to tool theft in
    2016 alone [1], strongly indicating the need for a robust and effective 
    theft mitigation system. We propose the BareTag Tool Tracker, a novel
    approach to tool tracking that utilizes Ultra-Wideband(UWB) and Bluetooth 
    Low-Energy(BLE) radio in order to real-time track tools, materials, or other 
    valuable items on a construction site. The system utilizes a series
    of pre-placed Anchor posts, that send UWB pings to a Tag that is connected
    to a tool. Each Anchor can then calculate its distance to a Tag, relaying
    that information to a Base station over Long Range(LoRa) radio. The 
    base station runs the aggregated distance data through a multilateration
    algorithm that can calculate the Tag's location with 50 cm accuracy.
    The calculated location is then output to a local terminal, as well as 
    uploaded to a cloud database for future reference. Altogether, the BareTag Tool-Tracker
    is highly accurate (50 cm), low-power (1 year of battery life), and
    scalable (increase range by adding additional Anchors).
\end{abstract}

@@

The specifications listed in Table 1 demonstrate the goals that our project 
aims to achieve by the time that we have a finalized product.
\begin{center}
\begin{table}
\caption{Design Goals and Testing Plan}
\begin{tabular}{|p{4cm}|p{4cm}|}
    \hline
    \textbf{Specification}  & \textbf{Testing Plan}\\
    \hline
    The system will track an item’s geo-coordinates with a reliability of 
    95\% within a specified 100 m x 100 m zone. 
    &
    Ping the Tag at randomized coordinates within the Sustainable 
    Engineering construction site.\\
    \hline
    The system will have less than 50 cm precision for tracking within 
    the specified zone.
    &
    Compare Tag’s measured location to its actual location 
    within a predefined grid.\\
    \hline
    The system will operate 24/7 for ~1 year without 
    replacing the batteries.
    &
    Measure the power consumption of the system in multiple modes.\\
    \hline
    Ensure that the Tag can be tracked via the Find My network outside 
    of the UWB zone.
    &
    Confirm tracking accuracy in scenarios with different saturations 
    of Find My devices by comparing with AirTags in the same location.\\
    \hline
    Able to increase range of the UWB perimeter by adding more Anchors.
    &
    Demonstrate that a Tag’s trackable range in the UWB zone increases 
    when adding a new Anchor.\\
    \hline
    Create a user-friendly GUI that displays geographical locations of multiple Tags.
    &
    Have a non-team member use our application and easily track Tags spread around a zone.\\
    \hline
\end{tabular}
\end{table}
\end{center}

@@

This device can be whatever system a supervisor desires to use as long as 
it can support a USB/Serial connection, or do so using adapters. The Host Device
performs the multilateration algorithms based on the data it receives from the LoRa Dongle.
The calculations get backhauled to a cloud database 
where the history of Tag locations can be mapped out to determine 
daily paths. This information is also displayed locally on a map of 
the desired area. The Host Device is also running the macless-haystack 
server, which acts as a means for tracking the Tags that have 
left the UWB area and are using BLE to integrate with the Apple Find 
My network. The Tags all have unique IDs that can be linked to tools or 
equipment, making it easy to parse the data and determine what items have 
been in what location. 
\subsection{3D Localization}
To find the z-coordinate of the Tag using only three Anchors, 
we need the distances between the Anchors and the distances between 
the Tag and Anchors. The shape the Anchors and Tag form is known as a 
tetrahedral. Using the volume and the base area of the tetrahedral we 
are able to find the height which represents the Tag’s z-coordinate. 
In [8] we know that using Heron’s formula on the distances between the 
Anchors we can get the area of the base
\[S_{ABC} = \sqrt{(p - AB)\cdot(p-AC)\cdot(p-BC)}\]
and using Euler’s formula we can get the volume of the tetrahedral.
\[V_{HABC}^2 = \frac{1}{36}
\begin{vmatrix}
    a^2 & \frac{a^2+b^2-n^2}{2} & \frac{a^2+c^2-m^2}{2} \\
    \frac{a^2+b^2-n^2}{2} & b^2 & \frac{b^2+c^2-l^2}{2} \\
    \frac{a^2+c^2-m^2}{2} & \frac{b^2+c^2-l^2}{2} & c^2
\end{vmatrix}\]

By rearranging the formula for finding the volume of a pyramid
\[V_{HABC} = \frac{1}{3}S_{ABC} \cdot H\]

we get the height of the tetrahedral as a function of volume and base.
\[H = 3 \cdot \frac{V_{HABC}}{S_{ABC}}\]
Then, using the volume and area that we previously calculated, 
we get the height of the tetrahedral, which is also the z-coordinate of
the Tag. To validate that this method works we ran a simulation through 
Blender and confirmed that the equations do indeed give us the correct 
z-coordinate.

\section{THE PROTOTYPE}
\subsection{Prototype Overview}
The prototype currently consists of 3 hardware components, each with 
associated software to control their functionality. The 3 hardware 
components are the Tag, the Anchor, and a LoRa dongle connected to the 
host computer.
The Tag consists of a DWM3001C UWB, BLE, and IMU combo module, along with 
associated programming headers, debug LEDs, and a 290 mAh LiPo battery. 
The DWM3001C allows us to use one module as our MCU and host device for 
all radios and other peripherals required for our design. The DWM3001C 
along with associated passive components are mounted on a custom SMT PCB 
to make the Tag. 
The Anchor also consists of a DWM3001C, as well as a RYLR998 LoRa radio, 
along with a programming header, debug LEDs, and a 2200 mAh LiIon battery. 
The RYLR998 is a hobbyist LoRa radio that allows us to transmit 256 byte 
packets wirelessly from our Anchors to the host computer. These packets 
communicate measured distance between Anchor and Tag, as well as data about 
the current state of the Anchor. Same as the Tag, the DWM3001C and RYLR998 
along with associated passive components are housed on a custom SMT PCB to 
make the Anchor. Anchors are mounted on metal poles that allow for fast 
reconfiguration and ease of testing indoors and outdoors. 
The LoRa dongle connected to the host computer is a generic USB TTL to 
UART converter connected to a RYLR998. The host computer runs a custom 
Python script that gathers distance data from all 3 Anchors to perform a 
trilateration algorithm to pinpoint the Tag’s current location. Anchors 
run a custom firmware that has them send a UWB ping to the Tag around once 
every second. A randomized delay is put between pings in order to prevent 
two different Anchors from pinging the Tag at that same time. Once receiving 
a distance measurement from the Tag, each Anchor relays this information to 
the host computer using the RYLR998, controlling the module over UART. The 
Tags run a simple custom firmware, responding to UWB pings whenever one is 
received. Combining these hardware and software components we get a complete 
prototype that can find a Tag’s location in real time with 50 cm accuracy.

\subsection{List of Hardware and Software}
Hardware:
\begin{itemize}
    \item Qorvo DWM3001C
    \item Reyax RYLR998
    \item 3.7V 290mAh LiPo
    \item 3.7 2200mAh LiIon
    \item USB-A to UART Converter
    \item Raspberry Pi 4B
    \item Tag-Connect Edge-Connect Programmer
    \item SEGGER JLink Edu Mini
\end{itemize}

Software:
\begin{itemize}
    \item Docker
    \item Xcode
    \item Serial Monitor
    \item Debian Linux
\end{itemize}

\subsection{Custom Hardware}
We designed two custom PCBs to act as the Tag and Anchors of our system. 
In our current system, we were able to trilaterally locate the custom PCB 
Tag with 3 Anchors: two Qorvo DWM3001CDK development boards and one of our 
custom PCB Anchors. Our main goal when creating our Tag PCB was to make it 
as small as possible. We eventually want these Tags to be placeable inside 
of tools. In our first revision, we were able to minimize the size of the Tag 
to approximately 26.8 by 31.6 mm squared. We were also able to successfully 
program our DWM3001C with the Edge-Connect seen in Fig 5 attached to the 
castellated edges of our Tag PCB. Because of this, we will be able to remove 
the 10-pin programming header seein in Fig. 3 to reduce the size of the Tag 
PCB in our second revision.

While the Tag and Anchor PCBs look quite different, they are populated with 
very similar components. The major difference between the two is the LoRa 
module that can be seen extending off the left side of the Anchor PCB in Fig. 
4. The LoRa module on the Anchor is used to communicate distance  measurements 
from the Anchors to a base station in order to perform our trilateration 
algorithm and locate the Tag.

While we were able to integrate our PCBs to make a functioning system, we 
still had some issues that we plan to address in our next revisions. The 
battery connectors into the PCBs on the Tags and the Anchors are facing 
into other components on the board, making it difficult to connect and 
disconnect our batteries to them. Our LEDs were meant to flash during 
transmission and reception of Ultra-Wideband pings on the Anchor and the 
Tag. While this was not achieved, the main functionality of our PCBs were 
not impacted. The UWB antenna on the Anchor PCB is on the opposite side of 
the LoRa antenna because we were originally worried about signals interfering 
with each other from the two different modules. In our next revision, we will 
have both antennas facing up to have them in their ideal orientation for 
transmitting and receiving signals. With this modification, we will be able to 
maximize the area where we can precisely locate our Tag with Ultra-Wideband 
pings. As you can see in Fig. 3 and Fig. 4, there are many more components on 
each of our boards. These include resistors to limit current flow into certain 
components, capacitors to reduce noise from input voltage, and LEDs that were 
meant to be used to help us debug our system. Most importantly, we had a 
voltage regulator to limit the input voltage into our Qorvo module to its 
preferred 3.3 volts. We also have many test points that helped us debug our 
system which we will remove in future revisions. Overall, our first revisions 
of our custom PCBs worked well enough for our prototype to function properly.
